\vspace*{-1.5em} \rule{\textwidth}{0.4pt} \\[0.4em] \textbf{ZKLoRA: Efficient Zero-Knowledge Proofs for LoRA Verification} \\[0.4em] \rule{\textwidth}{0.4pt}

### 2025-01-19 paper/main.tex
readme
update paper from main
polish Figure 2; proofread through section 2

\title{\textbf{ZKLoRA: Efficient Zero-Knowledge Proofs for LoRA Verification}}
\author{ \includegraphics[height=8em]{figs/bagel-logo-bw.png} \\
    \textbf{Bidhan Roy, Peter Potash, Marcos Villagra} \\
    Bagel Research Team\footnotemark[1] \\
}
\date{\today}

@@

  We present \textbf{ZKLoRA}, a zero-knowledge verification protocol that relies on polynomial commitments, succinct proofs, and multi-party inference to verify LoRA–base model compatibility without exposing LoRA weights. ZKLoRA produces \emph{deterministic} correctness guarantees and validates each LoRA module in \textbf{only 1--2 seconds} on statge of the art large language models. This low-latency approach enables nearly real-time verification and promotes secure collaboration among geographically dispersed teams and contract-based training pipelines. The protocol ensures that the delivered LoRA module works as claimed, safeguarding the contributor’s intellectual property while providing the base model user with verification of compatibility and lineage.
\end{abstract}

%------------------------------------------
% INTRO: emphasize verification challenge & clarify it's per-module
%------------------------------------------
\section{Introduction}
Large Language Models (LLMs) have attained remarkable success \cite{brown2020language, devlin2018bert}, but verifying fine-tuned modifications such as LoRA \cite{hu2021lora} in an untrusted, distributed training environment can be difficult when the updated weights must remain private. 
Traditionally, one might re-run an entire forward pass or inspect thousands of parameters to ensure correctness, which is infeasible for massive models. 
\texttt{ZKLoRA} addresses this by generating a zero-knowledge proof of correctness for each LoRA module, guaranteeing that the private LoRA genuinely fits the base model. 
Crucially, \textbf{the verification stage for each LoRA module} in \texttt{ZKLoRA} remains about 1--2 seconds, even at scales of multi-billion parameter, state of the art large language base models.

@@

We benchmarked \texttt{ZKLoRA} across several LLMs and smaller models with different numbers of LoRA modules. 
Our central question is how verification times, as well as settings and proof generation times, grow with \textbf{the number of LoRA modules}, while also considering each LoRA's average parameter size. 
Figure~\ref{fig:verify_vs_numlora} and Table~\ref{tab:settings_proof} detail this trade-off. \footnote{Note that the number of LoRA modules in a given model is not purely a function of the number of layers -- it is also a choice of which weight matrices within each layer is targeted. For example, just targeting one matrix per layer (ie the Query matrix in the Attention Block) will yeild one LoRA per layer. Conversely, targeting the Query, Key, and Value matrices yields three matrices per layer and makes the total number of LoRAs $3 \times num$\_$layers$.}

@@

Table~\ref{tab:settings_proof} similarly shows that \emph{proof generation} and \emph{settings time} scale with the number of LoRA modules, but also depend heavily on each module’s average size (indicated by larger or smaller dot size). These two steps (proof generation on the LoRA contributor’s side and cryptographic circuit setup) can become more expensive, yet remain feasible in decentralized settings or paid contract relationships. The \emph{Base Model User}, meanwhile, benefits from the relatively short verification overhead.

Overall, these plots confirm that \texttt{ZKLoRA} can handle multiple LoRA modules with minimal overhead for verifying correctness, emphasizing the viability of repeated or multi-adapter scenarios in large-scale LLM pipelines.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
    font=\small,
    >=latex,
    thick,
    node distance=1.6cm,
    stepbox/.style={
      draw,
      rounded corners,
      minimum width=6.2cm,
      minimum height=2.4cm,
      align=center,
      fill=gray!10,
      font=\bfseries
    },
    subblock/.style={
      draw,
      rounded corners,
      align=center,
      minimum width=1.6cm,
      minimum height=0.8cm,
      fill=gray!20,
      font=\scriptsize
    },
    arrow/.style={->, thick},
]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Step 1 BOX (Multi-Party Inference)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\node[stepbox,
      label={[above=2pt]:{\footnotesize \textbf{Step 1: Multi-Party Inference}}}] (step1) 
{\vspace{4pt}};  % minimal text inside, slight vertical spacing

% Inside Step 1: Two sub-blocks horizontally
\node[subblock,
      label={[font=\scriptsize]above:{Base Model User}},
      anchor=center, 
      xshift=-1.6cm,
      yshift=-0.4cm] 
(bmuser) at (step1.center)
{};

\node[subblock,
      label={[font=\scriptsize]above:{LoRA Contributor}},
      anchor=center, 
      xshift=+1.6cm,
      yshift=-0.4cm] 
(lorablk) at (step1.center)
{};

% Two separate arrows: "Base Acts" & "LoRA Acts"
\draw[arrow] 
  ([yshift=-0.05cm]bmuser.east) 
    to[bend left=20] 
  node[midway, above, yshift=-0.1cm, font=\scriptsize]{Base Acts}
  ([yshift=-0.05cm]lorablk.west);

\draw[arrow] 
  ([yshift=-0.25cm]lorablk.west) 
    to[bend right=20] 
  node[midway, below, yshift=+0.1cm, font=\scriptsize]{LoRA Acts}
  ([yshift=-0.25cm]bmuser.east);

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Step 2 BOX (Proof Generation), below step1
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\node[stepbox, fill=orange!15, 
      label={[above=2pt]:{\footnotesize \textbf{Step 2: Proof Generation}}}, 
      below=of step1] (step2)
{\footnotesize LoRA Contributor \\
Builds Zero-Knowledge Proofs};

% Arrow from step1 -> step2 with text "LoRA input/output acts"
\draw[arrow] (step1.south) -- node[left, xshift=-0.1cm, font=\scriptsize]{LoRA input/output acts} (step2.north);

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Step 3 BOX (Verification), below step2
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\node[stepbox, fill=blue!10, 
      label={[above=2pt]:{\footnotesize \textbf{Step 3: Verification}}}, 
      below=of step2] (step3)
{\footnotesize Base Model User \\
Verifies each proof};

% Arrow from step2 -> step3
\draw[arrow] (step2.south) -- node[left, font=\scriptsize]{\textit{proof files}} (step3.north);

\end{tikzpicture}
\caption{\textbf{Three-Step ZKLoRA Process (Vertical).}
\textbf{(1)} Base Model User and LoRA Contributor exchange ``Base Acts'' and ``LoRA Acts'' in a full multi-party inference. 
\textbf{(2)} The LoRA Contributor generates cryptographic proofs for correctness. 
\textbf{(3)} The Base Model User verifies these proofs, ensuring correct LoRA alignment without revealing private adapter weights.}
\label{fig:zklora_3steps}
\end{figure}

>>>>

\title{
    \vspace*{-1.5em}
    \rule{\textwidth}{0.4pt} \\[0.4em]
    \textbf{ZKLoRA: Efficient Zero-Knowledge Proofs for LoRA Verification} \\[0.4em]
    \rule{\textwidth}{0.4pt}
}
\author{ \includegraphics[height=8em]{figs/bagel-logo-bw.png} \\
    \textbf{Bidhan Roy, Peter Potash, Marcos Villagra} \\
    Bagel Research Team\footnotemark[1] \\
}
\date{\today}

@@

  We present \textbf{ZKLoRA}, a zero-knowledge verification protocol that relies on succinct proofs and multi-party inference to verify LoRA–base model compatibility without exposing LoRA weights. ZKLoRA produces \emph{deterministic} correctness guarantees and validates each LoRA module in \textbf{only 1--2 seconds} on state-of-the-art large language models. This low-latency approach enables nearly real-time verification and promotes secure collaboration among geographically dispersed teams and contract-based training pipelines. The protocol ensures that the delivered LoRA module works as claimed, safeguarding the contributor’s intellectual property while providing the base model user with verification of compatibility and lineage.
\end{abstract}

%------------------------------------------
% INTRO: emphasize verification challenge & clarify it's per-module
%------------------------------------------
\section{Introduction}
Large Language Models (LLMs) have attained remarkable success \cite{brown2020language, devlin2018bert}, but verifying fine-tuned modifications such as LoRA \cite{hu2021lora} in an untrusted, distributed training environment can be difficult when the updated weights must remain private. 
Traditionally, one might re-run an entire forward pass or inspect thousands of parameters to ensure correctness, which is infeasible for massive models. 
\texttt{ZKLoRA} addresses this by generating a zero-knowledge proof of correctness for each LoRA module, guaranteeing that the private LoRA genuinely fits the base model. 
Crucially, \textbf{the verification stage for each LoRA module} in \texttt{ZKLoRA} remains about 1--2 seconds, even at scales of multi-billion parameter, state-of-the-art large language base models.

@@

We benchmarked \texttt{ZKLoRA} across several LLMs and smaller models with different numbers of LoRA modules. 
Our central question is how verification times, as well as settings and proof generation times, grow with \textbf{the number of LoRA modules}, while also considering each LoRA's average parameter size. 
Figure~\ref{fig:verify_vs_numlora} and Table~\ref{tab:settings_proof} detail this trade-off.\footnotemark[2]

\footnotetext[2]{Note that the number of LoRA modules in a given model is not purely a function of the number of layers -- it is also a choice of which weight matrices within each layer is targeted. For example, just targeting one matrix per layer (ie the Query matrix in the Attention Block) will yeild one LoRA per layer. Conversely, targeting the Query, Key, and Value matrices yields three matrices per layer and makes the total number of LoRAs $3 \times num$\_$layers$.}

@@

Table~\ref{tab:settings_proof} similarly shows that average \emph{proof generation} and \emph{settings time} scale with the size of a LoRA module, which, combined with the number of modules, gives the total times. These two steps (proof generation on the LoRA contributor’s side and cryptographic circuit setup) can become more expensive, yet remain feasible in decentralized settings or paid contract relationships. The \emph{Base Model User}, meanwhile, benefits from the relatively short verification overhead.

Overall, these plots confirm that \texttt{ZKLoRA} can handle large-scale implementation of LoRA modules with minimal overhead for verifying correctness, emphasizing the viability of repeated or multi-adapter scenarios in large-scale LLM pipelines.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
    font=\small,
    >=latex,
    thick,
    node distance=1.6cm,
    stepbox/.style={
      draw,
      rounded corners,
      minimum width=6.2cm,
      minimum height=3.2cm,
      align=center,
      fill=gray!10,
      font=\bfseries
    },
    subblock/.style={
      draw,
      rounded corners,
      align=center,
      minimum width=1.6cm,
      minimum height=0.8cm,
      fill=gray!20,
      font=\scriptsize
    },
    arrow/.style={->, thick},
    % Define a style for the vertical arrow labels
    arrowlabel/.style={
      font=\footnotesize\sffamily\bfseries,
      fill=white,
      text=black,
      inner sep=1pt
    }
]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Step 1 BOX (Multi-Party Inference)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\node[stepbox] (step1) {%
  \footnotesize \textbf{Step 1: Multi-Party Inference}\\[6pt]
  \vspace{10pt}
};

% Sub-blocks within Step 1
\node[subblock,
      label={[font=\scriptsize]above:{Base Model User}},
      anchor=center, 
      xshift=-1.6cm,
      yshift=-1.0cm] (bmuser) at (step1.center)
{};

\node[subblock,
      label={[font=\scriptsize]above:{LoRA Contributor}},
      anchor=center, 
      xshift=+1.6cm,
      yshift=-1.0cm] (lorablk) at (step1.center)
{};

% Horizontal arrows in Step 1 (unchanged)
\draw[arrow] 
  ([yshift=-0.05cm]bmuser.east) 
    to[bend left=20] 
  node[midway, above, yshift=-0.1cm, font=\scriptsize]{Base Acts}
  ([yshift=-0.05cm]lorablk.west);

\draw[arrow] 
  ([yshift=-0.25cm]lorablk.west) 
    to[bend right=20] 
  node[midway, below, yshift=-0.05cm, font=\scriptsize]{LoRA Acts}
  ([yshift=-0.25cm]bmuser.east);

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Step 2 BOX (Proof Generation)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\node[stepbox, fill=orange!15, below=of step1] (step2) {%
  \footnotesize \textbf{Step 2: Proof Generation}\\[6pt]
  \footnotesize LoRA Contributor\\
  Builds Zero-Knowledge Proofs
};

% Vertical arrow from Step 1 to Step 2 with new style
\draw[arrow, shorten >=8pt, shorten <=8pt] (step1.south)
  -- node[arrowlabel, left, xshift=-0.1cm]{LoRA input/output acts}
  (step2.north);

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Step 3 BOX (Verification)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\node[stepbox, fill=blue!10, below=of step2] (step3) {%
  \footnotesize \textbf{Step 3: Verification}\\[6pt]
  \footnotesize Base Model User\\
  Verifies each proof
};

% Vertical arrow from Step 2 to Step 3 with new style
\draw[arrow, shorten >=8pt, shorten <=8pt] (step2.south)
  -- node[arrowlabel, left, xshift=-0.1cm]{proof files}
  (step3.north);

\end{tikzpicture}
\caption{\textbf{Three-Step ZKLoRA Process (Vertical).}
\textbf{(1)} Base Model User and LoRA Contributor exchange “Base Acts” and “LoRA Acts” in a multi-party inference.
\textbf{(2)} The LoRA Contributor generates cryptographic proofs for correctness.
\textbf{(3)} The Base Model User verifies these proofs, ensuring correct LoRA alignment without revealing private adapter weights.}
\label{fig:zklora_3steps}
\end{figure}
